Weighted Model Counting with Conditional Measures

### 2020-07-26 paper.tex
Mostly editing

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[capitalise]{cleveref}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usepackage[backgroundcolor=lightgray]{todonotes}
\usepackage{complexity}
\usepackage{soul}
\usepackage{subcaption}
\usepackage{multirow}

@@

\begin{itemize}
\item The Main Narrative
  \begin{enumerate}
  \item When weights are defined on literals, the measure on the free BA is
    fully independent.
  \item This means that the BA itself must be larger (i.e., have additional
    `meaningless' literals) to turn any probability distribution into an
    independent one.
  \item We show how we can define conditional weights on literals, allowing us
    to encode any probability distribution into a Boolean algebra that's not
    necessarily independent and thus can be smaller.
  \item We demonstrate a specific example of this by presenting a new way to
    encode Bayesian networks into instances of WMC and adapting a WMC algorithm
    (ADDMC) to run on the new format.
  \item We show that this results in significantly faster inference.
  \item We show that our encoding results in asymptotically fewer literals and
    fewer ADDs, and thus a simpler problem.
  \item (Maybe) we experimentally demonstrate a phase transition based on the
    number of variables per ADD.
  \end{enumerate}
\item Potential criticism may be that this is a step backwards and doesn't allow
  us to use SAT-based techniques for probabilistic inference. However, they can
  still be used for the 'theory+query' part.
  \begin{itemize}
  \item Zero-probability weights and one-probability weights can be interpreted
    as logical clauses. This doesn't affect ADDMC but could be useful for other
    solvers.
  \end{itemize}
\item[F] What are the main claims, what are the main takeaways, intuitive [???]
  of theorems to follow. To do this, we appeal to algebraic constructions to
  define the main concepts for introducing measures on Boolean algebras.
\item
  Algorithms\footnote{\url{http://beyondnp.org/pages/solvers/model-counters-exact/}}
  \begin{itemize}
  \item ADDMC \cite{DBLP:conf/aaai/DudekPV20} (rediscovered the
    multiplicativity of BAs in different words) (with optimal settings)
  \item Cachet \cite{DBLP:conf/sat/SangBBKP04}
  \item c2d \cite{DBLP:conf/ecai/Darwiche04}
  \item d4 \cite{DBLP:conf/ijcai/LagniezM17} (closed source, boo!)
  \item miniC2D  \cite{DBLP:conf/ijcai/OztokD15}
  \end{itemize}
\item Notable previous/related work
  \begin{itemize}
  \item Hailperin's approach to probability logic
    \cite{DBLP:journals/ndjfl/Hailperin84}
  \item Nilsson's (somewhat successful) probabilistic logic
    \cite{DBLP:journals/ai/Nilsson86,DBLP:journals/ai/Nilsson93}
  \item Logical induction: a big paper with a good overview of previous attempts
    to assign probabilities to logical sentences in a sensible way
    \cite{DBLP:journals/eccc/GarrabrantBCST16}
  \item Measures on Boolean algebras
    \begin{itemize}
    \item On possibility and probability measures in finite Boolean algebras
      \cite{DBLP:journals/soco/CastineiraCT02}
    \item Representation of conditional probability measures
      \cite{krauss1968representation}
    \end{itemize}
  \end{itemize}
\item Intuitively, a measure is just like a probability, except it's in
  $\mathbb{R}_{\ge 0}$ instead of $[0, 1]$.
\end{itemize}

@@

\begin{itemize}
\item[F] Give a concrete example of something impossible to represent using WMC.
\item[F] Can you say something here about factorized vs non-factorized
  weight function definitions? That is, factorized is when $w$ maps literals to
  $R_{\ge 0}$, non-factorized is when $w$ maps models to $R_{\ge 0}$ and
  \begin{itemize}
  \item come up with nice example when non-factorized weights are intuitive;
  \end{itemize}
\end{itemize}

@@

\todo[inline]{Provide a reference (and an explanation) for d-DNNF.}
\todo[inline]{If I'm citing Cachet, then maybe I should also mention what
  compilation tools they're using in their experiments.}

Hitherto, four techniques have been proposed for encoding Bayesian networks into
instances of WMC. We will identify them based on the initials of authors as well
as publications years: \texttt{d02} \cite{DBLP:conf/kr/Darwiche02},
\texttt{sbk05} \cite{DBLP:conf/aaai/SangBK05}, \texttt{cd05}
\cite{DBLP:conf/ijcai/ChaviraD05}, and \texttt{cd06}
\cite{DBLP:conf/sat/ChaviraD06}. Below we friendly summarise the observed
performance differences among them. Sang et al. \cite{DBLP:conf/aaai/SangBK05}
claim that \texttt{sbk05} is a smaller encoding than \texttt{d02} with respect
to both the number of clauses and the number of variables but provide no
experimental comparison. Chavira and Darwiche \cite{DBLP:conf/ijcai/ChaviraD05}
compare \texttt{cd05} with \texttt{d02} by measuring the time it takes to
compile either encoding into an arithmetic circuit (but do not measure inference
time). The more recent encoding \texttt{cd05} always compiles faster and results
in a smaller arithmetic circuit (as measured by the number of edges). In their
subsequent paper, the same authors perform two sets of experiments (that are
relevant to this summary) \cite{DBLP:conf/sat/ChaviraD06}. First, they compile
\texttt{cd05} and \texttt{cd06} encodings into d-DNNF, measuring both
compilation time and numbers of edges in the d-DNNF diagram. The results are
mostly in favour of \texttt{cd06}. Second, they compare the inference time of
\texttt{sbk05} run with Cachet \cite{DBLP:conf/sat/SangBBKP04} with the compile
times of \texttt{cd05} and \texttt{cd06}, but only on five (types of) instances.
In these experiments, \texttt{cd06} is always faster than \texttt{cd05}, while
the comparison with \texttt{sbk05} is mixed. Sometimes \texttt{cd06} is orders
of magnitude faster than \texttt{sbk05}, sometimes slightly slower. The
performance difference between \texttt{sbk05} and \texttt{cd05} is even harder
to judge: \texttt{sbk05} is better on three out of five instances and worse on
the remaining two. Based on this description, one would expect \texttt{cd06} to
be faster than both \texttt{cd05} and \texttt{sbk05}, both of which should be
faster than \texttt{d02}. The experiments in \cref{sec:experiments}, however,
strongly disagree with this prediction, showing that the quality of an encoding
depends strongly on the underlying search algorithm or compilation technique.

@@

Let $V$ denote the set of random variables in a Bayesian network. For any random
variable $X \in V$, let $\mathrm{pa}(X)$ denote the set of parents of $X$ and
$\im X$ denote the set of possible values.

@@

@@

@@

\subsection{An Example}

\todo[inline]{Reformulate this as a running example throughout the section.}

The Bayesian network in \cref{fig:example_bn} has
\begin{align*}
  V &= \{ W, F, T \}, \\
  \mathrm{pa}(W) &= \emptyset, \\
  \mathrm{pa}(F) &= \mathrm{pa}(T) = \{ W \}, \\
  \im W &= \im F = \{ 0, 1 \}, \\
  \im T &= \{ l, m, h \}, \\
  E(W) &= \{ \lambda_{W=1} \}, \\
  E(F) &= \{ \lambda_{F=1} \}, \\
  E(T) &= \{ \lambda_{T=l}, \lambda_{T=m}, \lambda_{T=h} \}, \\
  E^*(W) &= \{ \lambda_{W=1} \}, \\
  E^*(F) &= \{ \lambda_{F=1}, \lambda_{W=1} \}, \\
  E^*(T) &= \{ \lambda_{T=l}, \lambda_{T=m}, \lambda_{T=h}, \lambda_{W=1} \}, \\
  \mathrm{CPT_W} &= 0.5[\lambda_{W=1}]+0.5\overline{[\lambda_{W=1}]} = 0.5 \cdot 1, \\
  \mathrm{CPT_F} &= 0.6[\lambda_{F=1}] \cdot [\lambda_{W=1}] + 0.4[\lambda_{F=0}] \cdot [\lambda_{W=1}] + 0.1[\lambda_{F=1}] \cdot [\lambda_{W=0}] + 0.9[\lambda_{F=0}] \cdot [\lambda_{W=0}] \\
    &= 0.6[\lambda_{F=1}] \cdot [\lambda_{W=1}] + 0.4\overline{[\lambda_{F=1}]} \cdot [\lambda_{W=1}] + 0.1[\lambda_{F=1}] \cdot \overline{[\lambda_{W=1}]} + 0.9\overline{[\lambda_{F=1}]} \cdot \overline{[\lambda_{W=1}]}, \\
  \mathrm{CPT_T} &= ([\lambda_{T=l}] + [\lambda_{T=m}] + [\lambda_{T=h}]) \cdot (\overline{[\lambda_{T=l}]} + \overline{[\lambda_{T=m}]}) \cdot (\overline{[\lambda_{T=l}]} + \overline{[\lambda_{T=h}]}) \cdot (\overline{[\lambda_{T=m}]} + \overline{[\lambda_{T=h}]}) \cdot (\dots),
\end{align*}
and can be represented in a textual format as
\[
  \begin{array}{l r r l l}
    \lambda\sb{T=l} &\lambda\sb{T=m} &\lambda\sb{T=h} & &0 \\
                    &-\lambda\sb{T=l} &-\lambda\sb{T=m} & &0 \\
                    &-\lambda\sb{T=l} &-\lambda\sb{T=h} & &0 \\
                    &-\lambda\sb{T=m} &-\lambda\sb{T=h} & &0 \\
    w &\lambda\sb{W=1} & &0.5 &0.5 \\
    w &\lambda\sb{F=1} &\lambda\sb{W=1} &0.6 &0.4 \\
    w &\lambda\sb{F=1} &-\lambda\sb{W=1} &0.1 &0.9 \\
    w &\lambda\sb{T=l} &\lambda\sb{W=1} &0.2 &1 \\
    w &\lambda\sb{T=m} &\lambda\sb{W=1} &0.4 &1 \\
    w &\lambda\sb{T=h} &\lambda\sb{W=1} &0.4 &1 \\
    w &\lambda\sb{T=l} &\lambda\sb{W=0} &0.6 &1 \\
    w &\lambda\sb{T=m} &\lambda\sb{W=0} &0.3 &1 \\
    w &\lambda\sb{T=h} &\lambda\sb{W=0} &0.1 &1
  \end{array}
\]
with each $\lambda$ replaced with a unique positive integer. This format is
based on the format used by the Cachet solver \cite{DBLP:conf/sat/SangBBKP04}
to encode WMC problems, which extends the DIMACS format for CNF formulas with
weight clauses. Subsequently, we extend it in two ways:
\begin{itemize}
\item a single weight clause now supports an arbitrary number of literals,
\item and each weight clause has two probabilities instead of one (i.e., we no
  longer assume that $\Pr(v) + \Pr(\neg v) = 1$ for all variables $v \in U$).
\end{itemize}
The way we use this encoding, it is always the case that either both
probabilities sum to one, or the second probability (i.e., the probability for
the complement of the variable) is equal to one.

@@

\todo[inline]{Even though a lot of papers don't do this, I want a reference and
  a bit of a (historical/contextual) description here.}

ADDMC constructs the \emph{Gaifman graph} of the input CNF formula as an aid for
the algorithm's heuristics. This graph has as vertices the variables of the
formula, and there is an edge between two variables $u$ and $v$ if there is a
clause in the formula that contains both $u$ and $v$. We extend this definition
to functions on Boolean algebras, i.e., the factors of $\phi$. For any pair of
distinct variables $u, v \in U$, we draw an edge between them in the Gaifman
graph if there is a function $A\colon 2^X \to \mathbb{R}_{\ge 0}$ that is a
factor of $\phi$ such that $u \in X$ and $v \in X$. For instance, a factor such
as $\mathrm{CPT}_X$ will enable edges between all distinct pairs of variables in
$E^*(X)$.

@@

\begin{itemize}
\item We don't compare 'compile times' because our encoding time is linear, so
  we would easily beat everyone else.
\item When the Bayesian network has an evidence file, we compute the probability
  of evidence. Otherwise, let $X$ denote the last-mentioned node in the Bayesian
  network. If $\mathsf{true}$ is a valid value of $X$, we compute the marginal
  probability of $X = \mathsf{true}$. Otherwise, we pick the first value of $X$
  and calculate its marginal probability. This applies to the Grid data set (as
  intended) and also to two instances of Plan Reconstruction and roughly half of
  the instances from 2004-PGM that have empty evidence files.
\item After the experiments are finished, note the processor, memory per thread,
  and add the following acknowledgment.
\item All other encodings are implemented in
  Ace~3.0\footnote{\url{http://reasoning.cs.ucla.edu/ace/}} and should be
  compiled with \texttt{-encodeOnly} (i.e., don't compile the CNF into an AC)
  and \texttt{-noEclause} (i.e., only use standard syntax) flags.
\item Datasets
  \begin{itemize}
  \item binary Bayesian networks from Sang et
    al.\footnote{\url{https://www.cs.rochester.edu/u/kautz/Cachet/}}
    \cite{DBLP:conf/aaai/SangBK05}
    \begin{itemize}
    \item Grid (networks) (ratio 75 means that 75\% of the nodes are
      deterministic),
    \item Plan recognition (problems),
    \item Deterministic quick medical reference (what do the numbers mean? the
      README doesn't say).
    \end{itemize}
  \item Bayesian networks available with Ace
    \begin{itemize}
    \item \texttt{2004-pgm} \cite{DBLP:journals/ijar/ChaviraDJ06} (binary)
    \item \texttt{2005-ijcai} \cite{DBLP:conf/ijcai/ChaviraD05}. The Genie/Smile
      files have their own citation data that I should probably extract. This is
      the only dataset that has some non-binary networks.
    \item \texttt{2006-ijar} \cite{DBLP:journals/ijar/ChaviraDJ06} (binary)
    \end{itemize}
  \end{itemize}
\item \texttt{cd05} relaxes the encoding so much that extra models become
  possible. They are supposed to be filtered out by the algorithm, but mine
  can't do that because it doesn't deal with models. Same for \texttt{cd06}
  because it's based on \texttt{cd05}.
\end{itemize}

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[capitalise]{cleveref}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usepackage[backgroundcolor=lightgray]{todonotes}
\usepackage{complexity}
\usepackage{soul}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage[binary-units]{siunitx}

@@

\begin{itemize}
\item The Main Narrative
  \begin{enumerate}
  \item When weights are defined on literals, the measure on the free BA is
    fully independent.
  \item This means that the BA itself must be larger (i.e., have additional
    `meaningless' literals) to turn any probability distribution into an
    independent one.
  \item We show how we can define conditional weights on literals, allowing us
    to encode any probability distribution into a Boolean algebra that's not
    necessarily independent and thus can be smaller.
  \item We demonstrate a specific example of this by presenting a new way to
    encode Bayesian networks into instances of WMC and adapting a WMC algorithm
    (ADDMC) to run on the new format.
  \item We show that this results in significantly faster inference.
  \item We show that our encoding results in asymptotically fewer literals and
    fewer ADDs, and thus a simpler problem.
  \item (Maybe) we experimentally demonstrate a phase transition based on the
    number of variables per ADD.
  \end{enumerate}
\item Potential criticism may be that this is a step backwards and doesn't allow
  us to use SAT-based techniques for probabilistic inference. However, they can
  still be used for the 'theory+query' part.
  \begin{itemize}
  \item Zero-probability weights and one-probability weights can be interpreted
    as logical clauses. This doesn't affect ADDMC but could be useful for other
    solvers.
  \end{itemize}
\item[F] What are the main claims, what are the main takeaways, intuitive [???]
  of theorems to follow. To do this, we appeal to algebraic constructions to
  define the main concepts for introducing measures on Boolean algebras.
\item
  Algorithms\footnote{\url{http://beyondnp.org/pages/solvers/model-counters-exact/}}
  \begin{itemize}
  \item ADDMC \cite{DBLP:conf/aaai/DudekPV20} (rediscovered the
    multiplicativity of BAs in different words) (with optimal settings)
  \item Cachet \cite{DBLP:conf/sat/SangBBKP04}
  \item c2d \cite{DBLP:conf/ecai/Darwiche04}
  \item d4 \cite{DBLP:conf/ijcai/LagniezM17} (closed source, boo!)
  \item miniC2D  \cite{DBLP:conf/ijcai/OztokD15}
  \end{itemize}
\item Notable previous/related work
  \begin{itemize}
  \item Hailperin's approach to probability logic
    \cite{DBLP:journals/ndjfl/Hailperin84}
  \item Nilsson's (somewhat successful) probabilistic logic
    \cite{DBLP:journals/ai/Nilsson86,DBLP:journals/ai/Nilsson93}
  \item Logical induction: a big paper with a good overview of previous attempts
    to assign probabilities to logical sentences in a sensible way
    \cite{DBLP:journals/eccc/GarrabrantBCST16}
  \item Measures on Boolean algebras
    \begin{itemize}
    \item On possibility and probability measures in finite Boolean algebras
      \cite{DBLP:journals/soco/CastineiraCT02}
    \item Representation of conditional probability measures
      \cite{krauss1968representation}
    \end{itemize}
  \end{itemize}
\item Intuitively, a measure is just like a probability, except it's in
  $\mathbb{R}_{\ge 0}$ instead of $[0, 1]$.
\item Somewhere: explain things line CNF, CPT, etc.
\end{itemize}

@@

\todo[inline,caption={}]{
  \begin{itemize}
  \item[F] Give a concrete example of something impossible to represent
    using WMC.
  \item[F] Can you say something here about factorized vs non-factorized weight
    function definitions? That is, factorized is when $w$ maps literals to
    $R_{\ge 0}$, non-factorized is when $w$ maps models to $R_{\ge 0}$ and
    \begin{itemize}
    \item come up with nice example when non-factorized weights are intuitive;
    \end{itemize}
  \end{itemize}
}

@@

Hitherto, four techniques have been proposed for encoding Bayesian networks into
instances of WMC. We will identify them based on the initials of authors as well
as publications years: \texttt{d02} \cite{DBLP:conf/kr/Darwiche02},
\texttt{sbk05} \cite{DBLP:conf/aaai/SangBK05}, \texttt{cd05}
\cite{DBLP:conf/ijcai/ChaviraD05}, and \texttt{cd06}
\cite{DBLP:conf/sat/ChaviraD06}. Below we friendly summarise the observed
performance differences among them. Sang et al. \cite{DBLP:conf/aaai/SangBK05}
claim that \texttt{sbk05} is a smaller encoding than \texttt{d02} with respect
to both the number of clauses and the number of variables but provide no
experimental comparison. Chavira and Darwiche \cite{DBLP:conf/ijcai/ChaviraD05}
compare \texttt{cd05} with \texttt{d02} by measuring the time it takes to
compile either encoding into an arithmetic circuit (but do not measure inference
time). The more recent encoding \texttt{cd05} always compiles faster and results
in a smaller arithmetic circuit (as measured by the number of edges). In their
subsequent paper, the same authors perform two sets of experiments (that are
relevant to this summary) \cite{DBLP:conf/sat/ChaviraD06}. First, they compile
\texttt{cd05} and \texttt{cd06} encodings into d-DNNF (i.e., deterministic
decomposable negation normal form \cite{DBLP:journals/jancl/Darwiche01}),
measuring both compilation time and numbers of edges in the d-DNNF diagram. The
results are mostly in favour of \texttt{cd06}. Second, they compare the
inference time of \texttt{sbk05} run with Cachet \cite{DBLP:conf/sat/SangBBKP04}
with the compile times of \texttt{cd05} and \texttt{cd06}, but only on five
(types of) instances. In these experiments, \texttt{cd06} is always faster than
\texttt{cd05}, while the comparison with \texttt{sbk05} is mixed. Sometimes
\texttt{cd06} is orders of magnitude faster than \texttt{sbk05}, sometimes
slightly slower. The performance difference between \texttt{sbk05} and
\texttt{cd05} is even harder to judge: \texttt{sbk05} is better on three out of
five instances and worse on the remaining two. Based on this description, one
would expect \texttt{cd06} to be faster than both \texttt{cd05} and
\texttt{sbk05}, both of which should be faster than \texttt{d02}. The
experiments in \cref{sec:experiments}, however, strongly disagree with this
prediction, showing that the quality of an encoding depends strongly on the
underlying search algorithm or compilation technique.

@@

Let $V$ denote the set of random variables in a Bayesian network. For any random
variable $X \in V$, let $\mathrm{pa}(X)$ denote the set of parents of $X$ and
$\im X$ denote the set of possible values. We will use the Bayesian network in
\cref{fig:example_bn} (loosely based on an example by Sang et al.
\cite{DBLP:conf/aaai/SangBK05}) as a running example throughout this section.
For the Bayesian network in \cref{fig:example_bn}, we have:
\begin{align*}
  V &= \{ W, F, T \}, \\
  \mathrm{pa}(W) &= \emptyset, \\
  \mathrm{pa}(F) &= \mathrm{pa}(T) = \{ W \}, \\
  \im W &= \im F = \{ 0, 1 \}, \\
  \im T &= \{ l, m, h \}.
\end{align*}

@@

\todo[inline]{Come up with a less repetitive wording.}
For the Bayesian network in \cref{fig:example_bn}, we have:
\begin{align*}
  E(W) &= \{ \lambda_{W=1} \}, \\
  E(F) &= \{ \lambda_{F=1} \}, \\
  E(T) &= \{ \lambda_{T=l}, \lambda_{T=m}, \lambda_{T=h} \}, \\
  E^*(W) &= \{ \lambda_{W=1} \}, \\
  E^*(F) &= \{ \lambda_{F=1}, \lambda_{W=1} \}, \\
  E^*(T) &= \{ \lambda_{T=l}, \lambda_{T=m}, \lambda_{T=h}, \lambda_{W=1} \}.
\end{align*}

@@

\todo[inline]{Ditto.}
For the Bayesian network in \cref{fig:example_bn}, we have:
\begin{align*}
  \mathrm{CPT_W} &= 0.5[\lambda_{W=1}]+0.5\overline{[\lambda_{W=1}]} = 0.5 \cdot 1, \\
  \mathrm{CPT_F} &= 0.6[\lambda_{F=1}] \cdot [\lambda_{W=1}] + 0.4[\lambda_{F=0}] \cdot [\lambda_{W=1}] + 0.1[\lambda_{F=1}] \cdot [\lambda_{W=0}] + 0.9[\lambda_{F=0}] \cdot [\lambda_{W=0}] \\
    &= 0.6[\lambda_{F=1}] \cdot [\lambda_{W=1}] + 0.4\overline{[\lambda_{F=1}]} \cdot [\lambda_{W=1}] + 0.1[\lambda_{F=1}] \cdot \overline{[\lambda_{W=1}]} + 0.9\overline{[\lambda_{F=1}]} \cdot \overline{[\lambda_{W=1}]}, \\
  \mathrm{CPT_T} &= ([\lambda_{T=l}] + [\lambda_{T=m}] + [\lambda_{T=h}]) \cdot (\overline{[\lambda_{T=l}]} + \overline{[\lambda_{T=m}]}) \cdot (\overline{[\lambda_{T=l}]} + \overline{[\lambda_{T=h}]}) \cdot (\overline{[\lambda_{T=m}]} + \overline{[\lambda_{T=h}]}) \cdot (\dots).
\end{align*}

@@

\subsection{Textual Representation}

The Bayesian network in \cref{fig:example_bn} can be represented in a textual
format as
\[
  \begin{array}{l r r l l}
    \lambda\sb{T=l} &\lambda\sb{T=m} &\lambda\sb{T=h} & &0 \\
                    &-\lambda\sb{T=l} &-\lambda\sb{T=m} & &0 \\
                    &-\lambda\sb{T=l} &-\lambda\sb{T=h} & &0 \\
                    &-\lambda\sb{T=m} &-\lambda\sb{T=h} & &0 \\
    w &\lambda\sb{W=1} & &0.5 &0.5 \\
    w &\lambda\sb{F=1} &\lambda\sb{W=1} &0.6 &0.4 \\
    w &\lambda\sb{F=1} &-\lambda\sb{W=1} &0.1 &0.9 \\
    w &\lambda\sb{T=l} &\lambda\sb{W=1} &0.2 &1 \\
    w &\lambda\sb{T=m} &\lambda\sb{W=1} &0.4 &1 \\
    w &\lambda\sb{T=h} &\lambda\sb{W=1} &0.4 &1 \\
    w &\lambda\sb{T=l} &\lambda\sb{W=0} &0.6 &1 \\
    w &\lambda\sb{T=m} &\lambda\sb{W=0} &0.3 &1 \\
    w &\lambda\sb{T=h} &\lambda\sb{W=0} &0.1 &1
  \end{array}
\]
with each $\lambda$ replaced with a unique positive integer. This format is
based on the format used by the Cachet solver \cite{DBLP:conf/sat/SangBBKP04}
to encode WMC problems, which extends the DIMACS format for CNF formulas with
weight clauses. Subsequently, we extend it in two ways:
\begin{itemize}
\item a single weight clause now supports an arbitrary number of literals,
\item and each weight clause has two probabilities instead of one (i.e., we no
  longer assume that $\Pr(v) + \Pr(\neg v) = 1$ for all variables $v \in U$).
\end{itemize}
The way we use this encoding, it is always the case that either both
probabilities sum to one, or the second probability (i.e., the probability for
the complement of the variable) is equal to one.

@@

ADDMC constructs the \emph{Gaifman graph} \cite{gaifman1982local} of the input
CNF formula as an aid for the algorithm's heuristics. This graph has as vertices
the variables of the formula, and there is an edge between two variables $u$ and
$v$ if there is a clause in the formula that contains both $u$ and $v$. We
extend this definition to functions on Boolean algebras, i.e., the factors of
$\phi$. For any pair of distinct variables $u, v \in U$, we draw an edge between
them in the Gaifman graph if there is a function $A\colon 2^X \to
\mathbb{R}_{\ge 0}$ that is a factor of $\phi$ such that $u \in X$ and $v \in
X$. For instance, a factor such as $\mathrm{CPT}_X$ will enable edges between
all distinct pairs of variables in $E^*(X)$.

@@

\begin{itemize}
\item We don't compare 'compile times' because our encoding time is linear, so
  we would easily beat everyone else.
\item When the Bayesian network has an evidence file, we compute the probability
  of evidence. Otherwise, let $X$ denote the last-mentioned node in the Bayesian
  network. If $\mathsf{true}$ is a valid value of $X$, we compute the marginal
  probability of $X = \mathsf{true}$. Otherwise, we pick the first value of $X$
  and calculate its marginal probability. This applies to the Grid data set (as
  intended) and also to two instances of Plan Reconstruction and roughly half of
  the instances from 2004-PGM that have empty evidence files.
\item After the experiments are finished, note the processor, memory per thread.
\item The experiments were run on Intel Xeon Gold 6138 processor with an
  \SI{8}{\giga\byte} memory limit.
\item I will call my encoding \texttt{db20}.
\item All other encodings are implemented in
  Ace~3.0\footnote{\url{http://reasoning.cs.ucla.edu/ace/}} and were compiled
  with \texttt{-encodeOnly} (i.e., don't compile the CNF into an AC) and
  \texttt{-noEclause} (i.e., only use standard syntax) flags.
\item Datasets
  \begin{itemize}
  \item binary Bayesian networks from Sang et
    al.\footnote{\url{https://www.cs.rochester.edu/u/kautz/Cachet/}}
    \cite{DBLP:conf/aaai/SangBK05}
    \begin{itemize}
    \item Grid (networks) (ratio 75 means that 75\% of the nodes are
      deterministic),
    \item Plan recognition (problems),
    \item Deterministic quick medical reference (what do the numbers mean? the
      README doesn't say).
    \end{itemize}
  \item Bayesian networks available with Ace
    \begin{itemize}
    \item \texttt{2004-pgm} \cite{DBLP:journals/ijar/ChaviraDJ06} (binary)
    \item \texttt{2005-ijcai} \cite{DBLP:conf/ijcai/ChaviraD05}. The Genie/Smile
      files have their own citation data that I should probably extract. This is
      the only dataset that has some non-binary networks.
    \item \texttt{2006-ijar} \cite{DBLP:journals/ijar/ChaviraDJ06} (binary)
    \end{itemize}
  \end{itemize}
\item \texttt{cd05} relaxes the encoding so much that extra models become
  possible. They are supposed to be filtered out by the algorithm, but mine
  can't do that because it doesn't deal with models. Same for \texttt{cd06}
  because it's based on \texttt{cd05}.
\end{itemize}
